---
title: "data_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Analysis Notebook
Dylan Shulman, Katie Benzan, Marianeli De Leon

```{r}
# Loading libraries and cleaning data 

library(tidyverse)
library(janitor)
library(tidytext)

data_2020 <- read_csv("2020-08-19.csv") %>% clean_names %>% rename(row = x1)
data_2022 <- read_csv("2022-01-08.csv") %>% clean_names %>% rename(row = x1)
reviews_19m <- read_csv("bgg-19m-reviews.csv") %>% clean_names %>% rename(row = x1)
detailed_data <- read_csv("games_detailed_info.csv") %>% clean_names %>% rename(row = x1)
```
# Exploratory analysis 

```{r}
glimpse(detailed_data)

# Most high rated category

detailed_data %>% 
  filter(average > 7.5) %>% 
  group_by(boardgamecategory) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count)) %>% 
  head(10)

# User with the most reviews 

reviews_19m %>% 
  group_by(user) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  head(10)

# Does anyone call a game the "worst game?" Which is called that the most? Spoiler: it's Monopoly

reviews_19m %>% 
  filter(rating < 3) %>% 
  filter(str_detect(comment, "worst game")) %>% 
  group_by(name) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))

# Dataframe showing difference between 2022 reviews and 2020 reviews 

old_data <- data_2020 %>% 
  select(id, rank, average, bayes_average, users_rated) %>% 
  rename(rank_2020 = rank) %>% 
  rename(average_2020 = average) %>% 
  rename(bayes_average_2020 = bayes_average) %>% 
  rename(users_rated_2020 = users_rated)
pandy_data <- left_join(data_2022, old_data, by=("id"))
```

I think we'll have trouble discovering how many times visitors went to the website over the course of the pandemic because we don't have any date information from before 2020 as a control. The biggest limitation of the data is its lack of date and time for reviews. Otherwise, I believe we'll be able to answer our questions. 

## Question 1: Did the pandemic affect the number of reviews or the attitude of reviews?
```{r}
# First, gotta do a couple mutations to make difference columns 

pandy_data <- pandy_data %>% 
  mutate(rank_diff = rank - rank_2020) %>% 
  mutate(average_diff = average - average_2020) %>% 
  mutate(users_rated_diff = users_rated - users_rated_2020)

# Averages 

pandy_data %>% 
  filter(!is.na(average_diff)) %>% 
  summarize(total = sum(average_diff)) %>% 
  as.numeric / nrow(pandy_data) 

pandy_data %>% 
  filter(!is.na(users_rated_diff)) %>% 
  summarize(total = sum(users_rated_diff)) %>% 
  as.numeric / nrow(pandy_data)   
```
Average difference in rating: -0.02565801
Average number of new ratings: 124.5064

So with games gaining an average of 125 reviews and only an average change of -0.03, I don't think the pandemic had a significant impact on the dataset. However, averages may not be the best way to generalize a dataset of this size, as there could be plenty of outliers that could spawn off stories. 

```{r}
# New ratings 

users_22 <- pandy_data %>% 
  filter(!is.na(users_rated)) %>%
  summarize(total = sum(users_rated))

users_20 <- pandy_data %>% 
  filter(!is.na(users_rated_2020)) %>%
  summarize(total = sum(users_rated_2020))

users_22 - users_20
```
3,116,253 new ratings total. Across 21,831 games, that seems like a lot, but since we don't have another point from before the pandemic we can't establish whether or not this is comparitively an increase or decrease in ratings. 

```{r}
# Min and Max

max(pandy_data$average_diff, na.rm = TRUE) 
min(pandy_data$average_diff, na.rm = TRUE)
pandy_data %>% 
  filter(average_diff == 2.2)
pandy_data %>% 
  filter(average_diff == -3.51)
```
The minimum wasn't very interesting because it didn't have many ratings, but the max returned Pandemic Legacy board game, a 2020 release. It went up 2.2 points and got thousands more reviews over the two years. It's a game about a Soviet bioweapon. I found it curious that during the COVID-19 pandemic, the game that saw the greatest increase in rating is literally called Pandemic. I decided to do some more research into the individual reviews of Pandemic Legacy. 

```{r}
# Data set for exploration of Pandemic
pandemic_legacy_reviews <- reviews_19m %>% 
  filter(name == "Pandemic Legacy: Season 0")

# Some stats that may be enlightening 
median(pandemic_legacy_reviews$rating)
pandemic_legacy_reviews %>% 
  group_by(rating) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))

# How many new users rated the game? 
pandyusers <- data_2020 %>% 
  filter(name == "Pandemic Legacy: Season 0") %>% 
  select(users_rated) %>% 
  as.numeric

data_2022 %>% 
  filter(name == "Pandemic Legacy: Season 0") %>% 
  select(users_rated) %>% 
  as.numeric - pandyusers

# Did anyone mention COVID in their review? 
pandemic_legacy_reviews %>% 
  filter(str_detect(comment, "covid") | str_detect(comment, "COVID") | str_detect(comment, "Covid") | str_detect(comment, "corona") | str_detect(comment, "Corona"))

pandemic_legacy_reviews %>% 
  filter(!is.na(comment))
```
The game got significantly high reviews with a median of 9, meaning most of the reviews were 9s and 10s. It did however get a ton of new reviews, jumping from only 93 users rated to 3,477 users rated. This could explain the dramatic increase in rating over the two year period, but it's still surprised that this many users had a positive first experience. 

I thought it might be interesting to see if anyone even mentioned the global coronavirus pandemic, and I was surprised to return only two reviews. One did compare parts of the game to the COVID-19 pandemic, but the other one just said they couldn't play with their friends anymore because of the pandemic. I expected to see more reviewers correlating the two. Some of the reviews are in another language, and I didn't want to filter by "global pandemic" as I figured too many would be referencing the game and not real life, so it's conceivable that there are more reviews that mention COVID. 

## Question 1.5: Who's the biggest hater on the website?

With access to 19 million individual reviews, I was curious to see who was the most prolific reviewer who dished out the lowest ratings. I think it would be an interesting profile if there was a character that repeatedly dished out harsh critiques on a board game review website. 

I only wanted to deal with the reviewers in the top 1% of those with the most reviews, so that any small-time haters couldn't throw off the results. There are 412,816 users total, and only 4,131 in the top 1% of posting the most reviews. 

```{r}
# Create a dataset with the reviews from only reviewers in the top 1% of most reviews
counts <- reviews_19m %>% 
  group_by(user) %>% 
  summarize(review_count = n())
mostreviews <- quantile(counts$review_count, 0.99)
fulltime_reviewers <- reviews_19m %>% 
  left_join(counts, by="user") %>% 
  filter(review_count >= mostreviews)

# Create an empty dataframe to hold results
user_stats_all <- tibble()

# Duplicate the fulltime reviewers data set
uh_oh <- fulltime_reviewers %>% 
  filter(!is.na(rating))

# Grab names of users
names <- uh_oh %>% 
  group_by(user) %>% 
  summarize(count = n()) %>% 
  select(user)

# For loop
for(every_user in 1:nrow(names)) {

    # Keep only the row for a given row number, get rid of every other row
    user_name <- names %>%
      slice(every_user) %>% 
      as.character
      
    user_stats <- uh_oh %>% 
      filter(user == user_name) %>% 
      mutate(median_rating = median(rating)) %>% 
      mutate(max_rating = max(rating)) %>% 
      mutate(min_rating = min(rating)) %>% 
      mutate(mean_rating = mean(rating)) %>% 
      select(user, review_count, median_rating, max_rating, min_rating, mean_rating) %>% 
      distinct

    # Bind all rows
    user_stats_all <- user_stats_all %>%
      bind_rows(user_stats) %>% 
      distinct

}

# Display the completed dataframe
user_stats_all

# Make this a file so I don't have to run a 30 minute loop every time
write_csv(user_stats_all, "user_stats_all.csv")
```

```{r}
user_stats_all <- read_csv("user_stats_all.csv")
```

Now I have a nice and tidy dataset listing stats on all the users in the top 1% of those with the most reviews. Now I can do a simple arrange and find out if there are any users who make it their mission to leave negative reviews. 

```{r}
user_stats_all %>% 
  arrange(median_rating) 

user_stats_all %>% 
  arrange(mean_rating)
```

Two reviewers stick out: rabid_schnauzer and yahoo79. Both have mean ratings under 2, and both have medians of 1 nearing 1,000 reviews. I'll make datasets for both including only the reviews with comments to see if I can read any insight into these hypercritical critics. 

```{r}
shnauzer <- reviews_19m %>% 
  filter(user == "rabid_schnauzer" & !is.na(comment)) %>% 
  select(id, user, rating, comment, name) %>% 
  left_join(detailed_data, by=("id")) %>% 
  select(-row, -type, -primary, -thumbnail, -image, -suggested_num_players, -suggested_playerage, -suggested_language_dependence) 

shnauzer_all <- reviews_19m %>% 
  filter(user == "rabid_schnauzer") %>% 
  select(id, user, rating, comment, name) %>% 
  left_join(detailed_data, by=("id")) %>% 
  select(-row, -type, -primary, -thumbnail, -image, -suggested_num_players, -suggested_playerage, -suggested_language_dependence) 

yahoo <- reviews_19m %>% 
  filter(user == "yahoo79" & !is.na(comment))

yahoo_all <- reviews_19m %>% 
  filter(user == "yahoo79") %>% 
  select(id, user, rating, comment, name) %>% 
  left_join(detailed_data, by=("id")) %>% 
  select(-row, -type, -primary, -thumbnail, -image, -suggested_num_players, -suggested_playerage, -suggested_language_dependence) 

# What kind of games does shnauzer review?
shnauzer_all %>% 
  group_by(boardgamecategory) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))

shnauzer_all %>% 
  filter(str_detect(boardgamecategory, "Card Game") | str_detect(boardgamecategory, "Wargame"))

# What kind of games does yahoo review?
yahoo_all %>% 
  group_by(boardgamecategory) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))

yahoo_all %>% 
  filter(str_detect(boardgamecategory, "Card Game") | str_detect(boardgamecategory, "Wargame"))
```

rabid_schnauzer only gives us a view into their decidedly negative head 60 times, while yahoo79 never writes comments for their reviews. I wanted to see what kind of games rabid_schnauzer reviews, so I added on some details from the other set. Twenty-eight of their 60 reviews are of card games, just under half. In the full dataset, there are 381 reviews on card games or wargames out of the full 966 shnauzer reviews. 

yahoo also mostly reviews card games and wargames, making up 470 of his 1,144 reviews. Do the two haters review some of the same games? Do they agree in their rating? 

```{r}

ya <- yahoo_all %>% 
  select(id, name, rating, name) %>% 
  rename(yahoo_rating = rating)

haters <- shnauzer_all %>% 
  select(id, rating, comment) %>% 
  rename(shnauzer_rating = rating, shnauzer_comment = comment)

haters <- full_join(ya, haters, by="id")
haters_work_together <- haters %>% 
  filter(!is.na(shnauzer_rating)) %>% 
  filter(!is.na(yahoo_rating)) %>% 
  mutate(difference = case_when(
    yahoo_rating != shnauzer_rating ~ abs(yahoo_rating - shnauzer_rating)
  ))

haters_agree <- haters_work_together %>% 
  filter(is.na(difference)) %>% 
  arrange(desc(yahoo_rating))

haters_work_together
haters_disagree
```

The two critics have reviewed 417 of the same games, but agreed on rankings for 306 of those. The 306 common ratings are all 1/10s save for two ratings of three. Power Grid: Factory Manager and Cosmic Encounter were apparently good enough to warrant a solid 3/10 rating from these two haters. 

If they're so negative, what games did these critics actually give 10/10 ratings? 

```{r}
impressive <- haters %>% 
  filter((shnauzer_rating == 10) | (yahoo_rating == 10))

yahoo_all %>% 
  filter(rating == 10)

shnauzer_all %>% 
  filter(rating == 10)

haters_work_together %>% 
  filter((shnauzer_rating == 10) & (yahoo_rating == 10))
```
rabid_shnauzer has given 10/10 reviews significantly less than yahoo79, only dishing out 17 10s while yahoo had 80. There isn't a single game that both felt deserved a perfect score.  

Another question I have is, what exactly is shnauzer saying in his reviews with comments? Why doesn't this user like so many games?

```{r}
unique_words <- shnauzer %>% select(comment) %>%
  unnest_tokens(word, comment)

data("stop_words")

unique_words %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(25) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
      labs(x = "Count",
      y = "Unique words",
      title = "Count of unique words found in shnauzer reviews")
```

Looking at the unique words in shnauzer's comments wasn't very helpful, only bringing across generic game terms. Perhaps looking at bigrams will provide a bit more context. 

```{r}
bigrams <- shnauzer %>%
  unnest_tokens(bigram, comment, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  mutate(bigram = paste(word1, word2, sep=" ")) %>%
  group_by(bigram) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) 
bigrams %>%
  top_n(10)

bing <- get_sentiments("bing")

bing_word_counts <- unique_words %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE)
bing_word_counts
```
Interestingly enough, shnauzer actually uses positive words more often than negative words. Is this reviewer confused about the rating system? Also, the highest bigram is "hidden information." What games were they talking about?

```{r}
shnauzer %>% 
  filter(str_detect(comment, "hidden information"))

crux <- shnauzer %>% 
  filter(str_detect(comment, "rating"))
```
Looks like "hidden information" may be a component in some of the games shnauzer reviews. What's more interesting is what I found after I took a step back from the code and actually read some of shnauzer's reviews.  

After noticing the word "rating" strike so high on shnauzer's single-word rhetoric, I believe I've uncovered the crux of their negative mentality: they have a fundamental grudge against BBG's rating system. Just reading the comments of their reviews that mention "rating," this becomes glaringly apparent. In their reviews of 7 Wonders, shnauzer blames the popularity of the game on paid sponsorships, and so they take it upon themselves to "downvote as hard as I possibly can" and reverse this percieved capitalistic corruption of board game reviews. For Codex: Card-Time Strategy – Deluxe Set, shnauzer states "the BBG rating system is worse than Congress." 

This also explains why so much of shnauzer's diction is positive, since they often say that they thoroughly enjoyed a game but still felt it was overrated, necessitating a harsh rating to lower its status. 

However, shnauzer is no saint and is subject to immoral reviewing practices in their own right. For the game Dominion, they slashed the rating because they didn't like the fanbase of the game on the site.

But the best example is their review of Darkest Night: "A friend of mine makes money if you buy this, hence my rating is a rampant shill. Just like every single other rating on this site, except I give you an honest explanation." rabid_shnauzer is a lone righteous, hypocrite in a sea of thousands of board game reviewers. 

If I were pitching a story to an editor, although this doesn't classically fall into a newsworthy subject, I would use the rabid_shnauzer anecdote as a jumping point for a dive into this community. It sounds like there is a high level of user interaction on the site, and even some tension. It could make for a fun writeup to grab clicks online. 

Alternatively, it could be a starting point on a story that examines how exactly online reviews represent their products. It could span multiple websites, like Letterboxd and Goodreads, and decipher what the average ratings tell about a product: Is it sponsored? Viral? Or actually good quality? 

## Question 2: Do the most reviewed games have a trend of positive or negative reviews?
```{r}
highsusers <- quantile(data_2022$users_rated, 0.95)

most_rated <- data_2022 %>% 
  filter(users_rated >= highsusers) %>% 
  select(name, average, users_rated)

sum_head <- most_rated %>% 
  summarize(total = sum(average))
  
sum_average <- data_2022 %>% 
  summarize(total = sum(average))

sum_average / nrow(data_2022)
sum_head / nrow(most_rated)
```
Average rating across the entire dataset is 6.42, and average of the top 5% is 7.16. The most reviewed games have a higher average rating by over a full point than the average of all the games. Typically you might assume that people wopuld only review something online if they really loved it, or really hated it. It seems in this case the top 25 games were all loved by the majority. 

## Question 3: What type of game (multiplayer or single-player) has the most reviews? Most positive reviews? Most negative reviews?
```{r}
# most reviews
most_reviews <- reviews_19m %>% 
  group_by(name) %>% 
  summarise(
    number = n()
  ) %>% 
  arrange(desc(number))
```

Most reviews: Pandemic (108,971)

```{r}
# most positive reviews
# positive = rating > 7

pos_reviews <- reviews_19m %>% 
  filter(rating > 7) %>% 
  group_by(name) %>% 
  summarise(
    pos_number = n()
  ) %>% 
  arrange(desc(pos_number))
```

Most pos reviews: Pandemic (64,574)

```{r}
# most negative reviews
# negative = rating < 4

neg_reviews <- reviews_19m %>% 
  filter(rating < 4) %>% 
  group_by(name) %>% 
  summarise(
    neg_number = n()
  ) %>% 
  arrange(desc(neg_number))
```

Most neg reviews: Monopoly (10,457)

```{r}
# most neutral reviews
# neutral = 4 < rating < 7

neutral_reviews <- reviews_19m %>% 
  filter(rating < 7 & rating > 4) %>% 
  group_by(name) %>% 
  summarise(
    neutral_number = n()
  ) %>% 
  arrange(desc(neutral_number))
```

Most neutral reviews: Catan (28,966)

The game Pandemic seems to have the most positive reviews. Monopoly has the most negative reviews, and Catan has the most neutral reviews. It's ironic that during a pandemic, the board game became the most popular. This statistic insinuates that users were infatuated with the pandemic as much as they were locked down due to the virus. Contrary to presumption, Monopoly - a well-known, popular game - gathered the most negative reviews; however, about 10,000 reviews isn't close to the approximately 60,000 Pandemic gathered. The 10,000 reviews is the least of the top-3.

## Question 4: How did the recent pandemic changes affect the number of visits to the website?

```{r}
first_games_2020<-read_csv("data/2020-08-19.csv")%>% 
  clean_names %>% rename(row = x1)
second_games_2022<-read_csv("data/2022-01-08.csv")%>% 
  clean_names %>% rename(row = x1)
first_reviews_15m<-read_csv("data/bgg-15m-reviews.csv")%>% 
  clean_names %>% rename(row = x1)
second_reviews_19m<-read_csv("data/bgg-15m-reviews.csv")%>% 
  clean_names %>% rename(row = x1)
game_details<-read_csv("data/games_detailed_info.csv")%>% 
  clean_names %>% rename(row = x1)
```
```{r}
game_details %>% 
  filter(average > 7) %>% 
  group_by(boardgamecategory) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count)) %>% 
  head(10)
second_reviews_19m %>% 
  group_by(user) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  head(10)
second_reviews_19m %>% 
  filter(rating < 3) %>% 
  filter(str_detect(comment, "worst game")) %>% 
  group_by(name) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))
previous_games <- first_games_2020 %>% 
  select(id, rank, average, bayes_average, users_rated) %>% 
  rename(rank_2020 = rank) %>% 
  rename(average_2020 = average) %>% 
  rename(bayes_average_2020 = bayes_average) %>% 
  rename(users_rated_2020 = users_rated)
all_data <- left_join(second_games_2022, previous_games, by=("id"))
#How did the recent pandemic changes affect the number of visits to the website?
overall_average<-first_games_2020%>%
inner_join(second_games_2022, by=c("average"))
#This question can't be answered because the second file of games in 2022 on the website is significantly larger than the first ones in 2020. It would be impossible to remove the extra games from the second file in order to match the first. 
  View(mixed_games_average)
```

## Question 5: What are common phrases in the reviews? What about in positive reviews? Negative ones? 
```{r}
unique_words_comments <- reviews_19m %>%
  filter(rating==10)%>%
  filter(!is.na(comment)) %>% 
  select(comment) %>% 
  unnest_tokens(word, comment) %>%
  filter(!is.na(word))

unique_words_negative <- reviews_19m %>%
  filter(rating==1)%>%
  filter(!is.na(comment)) %>% 
  select(comment) %>% 
  unnest_tokens(word, comment) %>%
  filter(!is.na(word))
 
  unique_words_comments%>%
  count(word, sort = TRUE) %>%
  top_n(25) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
      labs(x = "Count",
      y = "Unique words",
      title = "Count of unique words found in reviews of Pandemic")
  
  data("stop_words")
  
  unique_words_comments %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(10)
```

```{r}
# Unfinished code for bigrams 
 unique_words_comments %>%
  unnest_tokens(bigram, word, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  mutate(bigram = paste(word1, word2, sep=" ")) %>%
  group_by(bigram) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(10)
```

