---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this my personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Working Directory

Above the file finder window, click "More" and select "Set as working directory." The command function setwd() will appear. 

## Libraries 

Install: install.packagaes()
Load: library()

Load tidyverse in pretty much every notebook. 

## Code Block

```{r}
# This is a comment. Code can run in these blocks. 
```

%>% means "and then..." adds a function
## Filter Functions

glimpse() get a sense of the data
group_by()
summarize() 
head() see first 6 rows of data
arrange(desc())

## Aggregate Functions IN PROGRESS FROM PRE LAB 1

summarise()

count_x = n()

total_x = sum(amount)
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))
