---
title: "Reverse Engineering Project"
author: "Dylan Shulman, Katie Benzan, Marianeli De Leon"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, "Over a Dozen Black and Latino Men Accused a Cop of Humiliating, Invasive Strip Searches. The NYPD Kept Promoting Him."
(https://www.propublica.org/article/over-a-dozen-black-and-latino-men-accused-a-cop-of-humiliating-invasive-strip-searches-the-nypd-kept-promoting-him)
`
## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
data <- read.csv("data/CCRB-Complaint-Data_202007271729/allegations_202007271729.csv")
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: 
"Eighty-six of the roughly 420 officers in the department who currently hold a rank above captain — running precincts and other large commands and overseeing hundreds of officers — have tallied at least one misconduct allegation that was substantiated by the CCRB, meaning that investigators amassed enough evidence of offenses, ranging from bad language to pistol whippings, to say that they happened and broke patrol guidelines. The most common involved improper property entries and searches."
* **Analysis summary**: 
I was not able to confirm the first part of the finding, since there are only 81 officers in the data set with a current rank above captain. The data includes only officers with substantiated claims and has three designations for officers above captain, "Chiefs and other ranks," "Deputy Inspector," and "Inspector." The reason I came up short could be because ProPublica "removed a small number of officers (62) against whom the CCRB had substantiated allegations, but whose substantiated allegations had not gone fully through the NYPD’s administrative prosecution process" from the data set. Perhaps the missing 5 officers were a part of the 62 removed from the data. However, I was able to confirm that "Premises entered and/or searched" was the most common allegation among this selection of substantiated claims. 

```{r}
# Put code to reverse engineer sentence here

# Claim 1, 86 out of 420 officers ranked above captain with substantiated complaints
data %>% 
  filter(rank_now == "Chiefs and other ranks" | rank_now == "Deputy Inspector" | rank_now == "Inspector") %>% 
  select(unique_mos_id) %>% 
  distinct() %>% 
  group_by(unique_mos_id) %>% 
  summarize(count = n()) %>% 
  summarize(total = sum(count))

# Claim 2, improper property entries and searches was the most common
data %>% 
  filter((rank_now == "Chiefs and other ranks" | rank_now == "Deputy Inspector" | rank_now == "Inspector") & str_detect(board_disposition, "^Subs")) %>% 
  group_by(allegation) %>% 
  summarize(count= n()) %>% 
  arrange(desc(count))

# Display results of code below this codeblock
```

### Sentence 2

* **Sentence text**: "Of at least 77 allegations made against him in 26 separate CCRB complaints, 29 were unsubstantiated; five were “unfounded,” meaning investigators concluded the incident never took place; and 27 were “exonerated,” meaning the conduct fell within patrol guidelines. Investigators have never proved that McCormack strip-searched men in the street, let alone penetrated them with his finger, though the city settled four lawsuits involving strip-search allegations against him for a total $184,500."
* **Analysis summary**: I was able to find almost every answer to this paragraph. I believe the data does not include the five unfounded complaints made against McCormack, and that's why our total complaint number is only 72 and the board_disposition column does not include an option of "unfounded." Otherwise, I was able to gather everything in the first sentence -- the second sentence was likely pulled from legal documents and figures not in this dataset, as there's no option for "lawsuit" or a number totaling money spent in lawsuits.  

```{r}
# Put code to reverse engineer sentence here
mccormack_complaints <- data %>% 
  filter(unique_mos_id == "18589")

#" Of at least 77 allegations..."
mccormack_complaints %>% nrow()

# "...in 26 separate CCRB complaints..."
mccormack_complaints %>% 
  group_by(complaint_id) %>% 
  summarise(count = n()) %>% 
  nrow()

# "...29 were unsubstantiated..."
mccormack_complaints %>% 
  filter(board_disposition == "Unsubstantiated") %>% 
  nrow()

# "...27 were “exonerated”..."
mccormack_complaints %>% 
  filter(board_disposition == "Exonerated") %>% 
  nrow()

# Display results of code below this codeblock
```

### Sentence 3

* **Sentence text**: “An overwhelming majority of the city’s 36,000 officers do not have a single substantiated complaint on their records; only 1 in 9 does.”
* **Analysis summary**: I was able to calculate this ratio with a simple equation after narrowing down the dataset to individual officers. With a clean dataset like this, it was a  straightforward process. 

```{r}
# Put code to reverse engineer sentence here

# finding the number of officers in the dataset 
total_officers <- data %>% 
  group_by(unique_mos_id) %>% 
  summarise(count = n()) %>% 
  nrow()

# the math to get the ratio 
total_officers / 36000 * 9

# Display results of code below this codeblock
```

### Sentence 4

* **Sentence text**: “Data shows that while more than half of officers in the lowest rank are people of color, the higher ranks — above captain — are more than 75% white. A dozen of those high-ranking officers have three or more complaints with substantiated allegations; 10 are white. Five have reached the rank of deputy inspector, the lowest promoted at the commissioner’s discretion; all are white.”
* **Analysis summary**: The first sentence uses a separate dataset from an article by The City, so we can't solve that from the ProPublica data. Unfortunately I wasn't able to find 12 high ranking officers with three or more complaints; only 5. Either my code is faulty or the reporters found the others with outside reporting. I did find that the five were all white, although only two were deputy inspectors. 

```{r}
# Put code to reverse engineer sentence here

# Claim 1, A dozen high ranking officers have three or more complaints
data %>% 
  filter((rank_now == "Chiefs and other ranks" | rank_now == "Deputy Inspector" | rank_now == "Inspector") & str_detect(board_disposition, "^Subs")) %>% 
  distinct(unique_mos_id, complaint_id) %>% 
  group_by(unique_mos_id) %>% 
  summarise(complaints = n()) %>% 
  filter(complaints >= 3) %>%
  nrow()

# Claim 2, majority white high officers with complaints 
data %>% 
  distinct(unique_mos_id, mos_ethnicity) %>% 
  filter(unique_mos_id == "18589" | unique_mos_id == "19489" | unique_mos_id == "19873" | unique_mos_id == "25814" | unique_mos_id == "25998") %>% 
  group_by(mos_ethnicity) %>% 
  summarize(count = n())

# Claim 3, five reached deputy inspector 
data %>% 
  distinct(unique_mos_id, rank_now) %>% 
  filter(unique_mos_id == "18589" | unique_mos_id == "19489" | unique_mos_id == "19873" | unique_mos_id == "25814" | unique_mos_id == "25998") %>% 
  group_by(rank_now) %>% 
  summarize(count = n())

# Display results of code below this codeblock
```

### Sentence 5

* **Sentence text**: “In all, the CCRB substantiated 16 allegations lodged in six complaints, most involving the stops and searches of young men of color. No other high-ranking NYPD officer has amassed as many substantiated CCRB complaints.”
* **Analysis summary**: We were able to fully confirm this paragraph. In this dataset, McCormack has 16 substantiated allegations in 16 rows. I narrowed these down to the six complaints, and found that most were stops and searches on men of color, however there were some NAs in the complainant ethnicity column. Finally, I searched the entire dataset to find that McCormack is the only officer with 6 substantiated complaints. 

```{r}
# Put code to reverse engineer sentence here

# Claim 1, McCormack had 16 substantiated allegations 
data %>% 
  filter(unique_mos_id == "18589" & str_detect(board_disposition, "^Subs")) %>% 
  nrow()

# Claim 2, Six complaints, most involving the stops and searches of young men of color
data %>% 
  filter(unique_mos_id == "18589" & str_detect(board_disposition, "^Subs")) %>% 
  distinct(complaint_id) %>% 
  nrow()
data %>% 
  filter(unique_mos_id == "18589" & str_detect(board_disposition, "^Subs")) %>% 
  distinct(allegation, complaint_id, complainant_ethnicity, complainant_gender) %>% 
  group_by(allegation, complainant_ethnicity, complainant_gender) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
# Claim 3, no other officer has this many
data %>% 
  filter((rank_now == "Chiefs and other ranks" | rank_now == "Deputy Inspector" | rank_now == "Inspector") & str_detect(board_disposition, "^Subs")) %>% 
  distinct(complaint_id, unique_mos_id) %>% 
  group_by(unique_mos_id) %>% 
  summarise(complaints = n()) %>%  
  filter(complaints >= 6)

# Display results of code below this codeblock
```

-30-